---
title: "HDT2"
output: html_document
date: '2022-08-01'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("forecast")
##install.packages("fUnitRoots")
#install.packages("ggfortify")

library(forecast)
library(tseries)
library(fUnitRoots)
library(ggfortify)
library(readxl)
getwd()
df_consumo <- read_excel("CONSUMO-2022-05.xlsx", range="B7:G276")
df_importacion <- read_excel("IMPORTACION-VOLUMEN-2022-05.xlsx", range="B7:G264")
```
```{r}
df_consumo
```

```{r}
df_importacion
```
```{r}
df_consumo$Diesel <- df_consumo$`Diesel alto azufre`+df_consumo$`Diesel bajo azufre`
df_consumo$`Diesel alto azufre` <- NULL
df_consumo$`Diesel bajo azufre` <- NULL
df_consumo[4] <- NULL
head(df_consumo)



summary(df_consumo)
plot(df_consumo$Fecha, df_consumo$`Gasolina superior`, xlab="FECHA", ylab="CONSUMO GASOLINA SUPERIOR", col= "blue")
points(df_consumo$Fecha, df_consumo$Diesel, col = "green", pch = 19)
points(df_consumo$Fecha, df_consumo$`Gasolina regular`, col = "red", pch = 19)


```
```{r}
df_importacion$Diesel <- df_importacion$`Diesel alto azufre`+df_importacion$`Diesel bajo azufre`
df_importacion$`Diesel alto azufre` <- NULL
df_importacion$`Diesel bajo azufre` <- NULL
df_importacion[4] <- NULL
head(df_importacion)
```
```{r}
library(tseries)
library(forecast)


Consumo_Diesel <- ts(df_consumo$Diesel, start=c(2000, 1), frequency = 12)
Consumo_regular <- ts(df_consumo$`Gasolina regular`, start=c(2000, 1), frequency = 12)
Consumo_super <- ts(df_consumo$`Gasolina superior`, start=c(2000, 1), frequency = 12)
importacion_Diesel <- ts(df_importacion$Diesel, start=c(2000, 1), frequency = 12)
importacion_regular <- ts(df_importacion$`Gasolina regular`, start=c(2000, 1), frequency = 12)
importacion_super <- ts(df_importacion$`Gasolina superior`, start=c(2000, 1), frequency = 12)


plot(Consumo_Diesel)
plot(Consumo_regular)
plot(Consumo_super)
plot(importacion_Diesel)
plot(importacion_regular)
plot(importacion_super)

